The competing effects of the immune system and EMT on epithelial cancers

### 2019-03-19 Paper/Paper.tex
Equation Updates
Changed parameter names as Adam and I discussed

$$ p_{prolif} = p(1+\delta_{\text{PM}}\text{prolif}_{\text{up}})(1-\delta_{MES}m_{GA})\frac{N_0}{N_0+N} $$

@@

$$ p_{apop} = d(1-\delta_{\text{AM}}\text{apop}_{\text{down}}) $$

@@

$$ p_{\text{NK}} =\delta_M \frac{\text{NK}}{N/N_{1}+\text{NK}}  \frac{E_{\text{NK}}}{1+\text{Treg}/\text{Treg}_{\text{EC50}}} (1-\delta_{IEM}\text{IE}_{\text{up}})(1-\delta_{MES}m_{\text{IE}}) $$
% Should I worry about the order in which these appear? Rearranging would make the formula look better, but the order I have introduced them makes the most sense.
% Also, any thoughts on NK, CTL, Treg not being italicized? While other capital letters are?

@@

$$ p_{rest} = 1 + \delta_{\text{MES}}p(1+\delta_{\text{PM}}\text{prolif}_{\text{up}})m_{GA}\frac{N_0}{N_0+N} $$

@@

%$$ \text{TGFB}_i = H(\text{TGFB}/N) + X_i, \quad X_i \sim N(0,\sigma^2)$$

$$ \text{TGFB}_i = \frac{\text{TGFB}_{max}}{N}\frac{\text{TGFB}/\gamma_{EC50}}{1+\text{TGFB}/\gamma_{EC50}}+ X_i
, \quad X_i \sim N(0,\sigma^2)$$
The summand on the left divides up the total TGF-$\beta$ among all $N$ cells evenly and a Hill function on this quantity determines how much TGF-$\beta$ each cell thus receives.
Then, some randomness is introduced by the independent and identically distributed variables $X_i$where $\sigma$ is a parameter in the model.
If this quantity is large enough, relative to the EMT score of the cell, then the cell will undergo EMT and its EMT score will increase.
If not, then the cell undergoes the reverse process, MET, and its EMT score will decrease.
Each cell then is relabeled as either epithelial or mesenchymal depending on its new EMT score and whether it is below or above the mesenchymal threshold.
Thus, there are two main factors that determine if a cell will end a cell cycle as mesenchymal: quantity of TGF-$\beta$ in the system and the current EMT score of the cell.

@@

$$ \text{TGFB} = \text{tgfb}_{\text{mut}}M + \text{tgfb}_{\text{Treg}}\text{Treg}$$ % clear that M is number of mutants, right?

@@

$$ \text{NK}' = \sigma_{\text{NK}} - d_{\text{NK}}\text{NK} $$

@@

$$ \text{NK}_{k+1} = \left (\text{NK}_k-\frac{\sigma_{\text{NK}}}{d_{\text{NK}}} \right )\exp(-d_{\text{NK}}\Delta t)+\frac{\sigma_{\text{NK}}}{d_{\text{NK}}} $$

%%%% Which of these two paragraphs is better? Currently the intro does not mention APCs and I don't think I need to go into that, especially when APC_k = \tilde{M}_k
For CTLs and Tregs, they rely on mutant cells being cleared before they can be activated.
Let $\tilde{M}_k$ represent the number of mutant cells cleared by the immune system during cell cycle $k$.
In addition, Treg recruitment is upregulated by TGF-$\beta$, which will be incorporated via a Hill function.
We choose the following differential equations to govern the CTL and Treg populations:

\begin{align*}
\text{CTL}' & = \sigma_{\text{CTL}}\tilde{M} - d_{\text{CTL}}\text{CTL} \\
\text{Treg}' & = \sigma_{\text{Treg}}\tilde{M} \frac{\text{TGFB}}{1+\text{TGFB}/\text{EC50}}- d_{\text{Treg}}\text{Treg}
\end{align*}

@@

\begin{align*}
\text{CTL}_{k+1} & = & \left (\text{CTL}_k-\sigma_{\text{CTL}}\tilde{M}_k/d_{\text{CTL}}\right )\exp(- d_{\text{CTL}}\Delta t) + \sigma_{\text{CTL}}\tilde{M}_k/d_{\text{CTL}}\\
\text{Treg}_{k+1} & = & \left (\text{Treg}_k-\frac{\sigma_{\text{Treg}}\tilde{M}_k}{d_{\text{Treg}}} \frac{\text{TGFB}}{1+\text{TGFB}/\text{EC50}}\right )\exp(-d_{\text{Treg}}\Delta t)+\\
& & \frac{\sigma_{\text{Treg}}\tilde{M}_k}{d_{\text{Treg}}} \frac{\text{TGFB}}{1+\text{TGFB}/\text{EC50}}
\end{align*}

>>>>

\begin{equation}\tag{2.1}
\rho_P = p(1+\delta_{P}\Delta_P)(1-\zeta \Delta_{MGA})\frac{K_0}{K_0+N_C}
\end{equation}

@@

\begin{equation}\tag{2.2}
\rho_A = d_C(1-\delta_{A}\Delta_A)
\end{equation}

@@

\begin{equation}\tag{2.3}
\rho_{NK} =\delta_{MUT} \frac{N_{NK}}{N_C/K_{1}+N_{NK}}  \frac{E_{NK}}{1+N_{Treg}/K_2} (1-\delta_{IE}\Delta_{IE})(1-\zeta \Delta_{MIE})
\end{equation}
% Should I worry about the order in which these appear? Rearranging would make the formula look better, but the order I have introduced them makes the most sense.
% Also, any thoughts on NK, CTL, Treg not being italicized? While other capital letters are?

@@

\begin{equation}\tag{2.4}
\rho_R = 1 + \zeta p(1+\delta_P\Delta_P)\Delta_{MGA}\frac{K_0}{K_0+N_C}
\end{equation}

@@

%$$ \text{TGFB}_i = H(\text{TGFB}/N) + X_i, \quad X_i \sim N(0,\sigma^2)$$
\begin{equation}\tag{2.5}
\tau_i = \frac{\tau_{max}}{N_C}\frac{1}{1+K_3/\tau}+ X_i
, \quad X_i \sim N(0,\sigma^2)
\end{equation}

The summand on the left divides up the total TGF-$\beta$ among all $N$ cells evenly and a Hill function on this quantity determines how much TGF-$\beta$ each cell thus receives.
Then, some randomness is introduced by the independent and identically distributed variables $X_i$where $\sigma$ is a parameter in the model.
If this quantity is large enough, relative to the EMT score of the cell, then the cell will undergo EMT and its EMT score will increase.
If not, then the cell undergoes the reverse process, MET, and its EMT score will decrease.
Each cell then is relabeled as either epithelial or mesenchymal depending on its new EMT score and whether it is below or above the mesenchymal threshold.
Thus, there are two main factors that determine if a cell will end a cell cycle as mesenchymal: quantity of TGF-$\beta$ in the system and the current EMT score of the cell.

@@

\begin{equation}\tag{2.6}
\tau = \tau_{MUT}N_{MUT} + \tau_{Treg}N_{Treg}
\end{equation}

@@

\begin{equation}\tag{2.7}
N_{NK}' = \sigma_{NK} - d_{NK}N_{NK}
\end{equation}

@@

 \begin{equation}\tag{2.8}
N_{\text{NK}}(k+1) = \left (N_{\text{NK}}(k)-\frac{\sigma_{\text{NK}}}{d_{\text{NK}}} \right )\exp(-d_{\text{NK}}\Delta t)+\frac{\sigma_{\text{NK}}}{d_{\text{NK}}}
\end{equation}

%%%% Which of these two paragraphs is better? Currently the intro does not mention APCs and I don't think I need to go into that, especially when APC_k = \tilde{M}_k
For CTLs and Tregs, they rely on mutant cells being cleared before they can be activated.
Let $N_{\text{MUT}}^*(k)$ represent the number of mutant cells cleared by the immune system during cell cycle $k$.
In addition, Treg recruitment is upregulated by TGF-$\beta$, which will be incorporated via a Hill function.
We choose the following differential equations to govern the CTL and Treg populations:

\begin{equation}\tag{2.9}
\begin{split}
N_\text{CTL}' & = \sigma_{\text{CTL}}N_{\text{MUT}}^* - d_{\text{CTL}}\text{CTL} \\
\text{Treg}' & = \sigma_{\text{Treg}}N_{\text{MUT}}^* \frac{\tau}{1+\tau/K_4}- d_{\text{Treg}}N_\text{Treg}
\end{split}
\end{equation}

@@

\begin{equation}\tag{2.10}
\begin{split}
N_\text{CTL}(k+1) & =  \left (N_\text{CTL}(k)-\sigma_{\text{CTL}}N_{\text{MUT}}^*(k)/d_{\text{CTL}}\right )\exp(- d_{\text{CTL}}\Delta t) + \sigma_{\text{CTL}}N_{\text{MUT}}^*(k)/d_{\text{CTL}}\\
N_\text{Treg}(k+1) & =  \left (N_\text{Treg}(k)-\frac{\sigma_{\text{Treg}}N_{\text{MUT}}^*(k)}{d_{\text{Treg}}} \frac{\tau}{1+\tau/K_4}\right )\exp(-d_{\text{Treg}}\Delta t)\\
&+ \frac{\sigma_{\text{Treg}}N_{\text{MUT}}^*(k)}{d_{\text{Treg}}} \frac{\tau}{1+\tau/K_4}
\end{split}
\end{equation}
